Global land cover fraction mapping: methods, variables, challenges

### 2020-04-29 article/article.tex
Changes from papeeria

\abstract{The current common practice for producing global land cover maps is to represent land cover classes as discrete units, e.g. a pixel may be labelled as grassland, tree cover, or open water. However, in reality, the area covered by the pixel is hardly ever homogeneous, it may be e.g. trees surrounded by grass next to water. This mismatch between the reality and its representation in maps causes a loss in precision, as only the dominant land cover class can be represented in such ``hard'' classification maps. This problem is worse for coarse resolution land cover maps. A common workaround to this issue is to define mosaic classes, i.e. land cover classes that are combinations of other land cover classes, such as open forest (combination of grassland and tree cover). However, this approach leads to increasingly complex map legends and confusion about the definitions.

Our proposed solution to this issue is to employ fractional land cover mapping: for each pixel, define the proportion of the pixel that each considered land cover class occupies. This allows expressing an open forest as e.g. 40\% grassland, 60\% tree cover, and 0\% of the rest of the land cover classes. This ``fuzzy'' classification not only allows to more precisely express the land cover that exists in reality, but also gives users the flexibility to derive their own maps with a customised legend. For instance, if a user defines open forest as one that has less than 50\% tree cover and at least 30\% of grassland, they can generate a map with this definition from land cover fraction data, rather than relying on the definition that the mapmakers provide.
Thus exhaustive land cover fraction data is desired by a number of user communities.

In this paper, we compared the performance of eight very machine learning algorithms for land cover fraction mapping: random forest regression, multilayer perceptron neural networks, partial least squares regression, fuzzy nearest prototype classification, multinomial logistic regression, lasso regression, cubist regression and support vector regression. In order to allow the algorithms to make good predictions about land cover fractions, six types of covariates were used: spectral bands and vegetation indices derived from the Proba-V 100 m TOC reflectance data; temporal features derived from the full time series of said imagery; terrestrial features from a digital elevation model; soil properties and climate data from global models, and location. We compared the contribution of each of the covariate types to improving the accuracy of the model predictions.

One challenge when dealing with fractional data is the data imbalance: in most locations, only a single or a few classes dominate a pixel, with the rest taking up 0\% of the pixel. Values at the extremes (0\% and 100\%, i.e. pure pixels) are difficult for machine learning algorithms to estimate. We propose a new solution to this problem by combining three models: two classification models and one regression model. The first classification model determines whether a pixel is pure, and if so, it is processed using a classification model, otherwise it is processed with a regression model.

% Validation
% - 28k points
% - Independent training and validation sets

The models were trained on 138164 training points throughout the globe. The training points, collected as part of the Copernicus Global Land System: Dynamic Land Cover (CGLS-LC100) project, describe the fraction that each land cover class (bare soil, crops, grassland, shrubs, trees, urban/built-up, water and wetlands) occupies within a 100\texttimes100 m area, which is aligned to the grid used for Proba-V 100 m products. After the models were trained, their land cover class fraction predictions were compared with an independent validation set of 20705 points, which are likewise aligned to the Proba-V grid. Overall as well as per-class RMSE, MAE, and ME statistics were used to compare the models. In addition, the subpixel confusion-uncertainty matrix was used to analyse which classes were the most difficult to separate for each model, as well as to derive overall, users' and producers' accuracies and the kappa statistic. The models were compared with each other, as well as against a control intercept-only model.

Results show that the three-model approach increases the overall accuracy of the predictions by more accurately modelling the extreme values. From the machine learning models that were compared, random forest and cubist regressions reaches the highest overall accuracy of 72\%. This is comparable to the reported accuracies of the current ``hard'' global land cover maps. Random Forest covariate importance analysis shows that spectral covariates are the most important for predicting land cover fractions, followed by the location, terrestrial and temporal covariates.

The application of the method on data gathered throughout the globe shows the applicability of the method to a variety of land cover conditions, ranging from highly heterogeneous agricultural landscapes to homogeneous bodies of water. This research proves that machine learning algorithms can be applied globally to create a global fractional land cover product. Such a product would cater to users with different fields of focus. Among other benefits, more precise estimation of land cover allows for 1) improved climate models to help reach Sustainable Development Goals (for which land cover is a key variable), 2) facilitate the making of models for vegetation and urban area dynamics, and 3) help governments and land owners keep better track of relevant land cover and its change. In addition, this work contributes to the operationalisation of the CGLS-LC100 project, which aims to produce exhaustive global land cover fractions as an operational product.

>>>>

\abstract{Currently most global land cover maps represent land cover classes as discrete units. However, in reality pixels often contain mixed land cover. This causes loss of precision as only the dominant land cover class may be represented. A common workaround to this issue is to define mosaic classes that are combinations of pure land cover classes mixed with certain ratios, e.g. open forest as a combination of grassland and trees. However, this approach leads to increasingly complex map legends and confusion about the definitions.

Our proposed solution is to map land cover fractions: define the proportion that each land cover class occupies within the pixel. Such fuzzy classification both increases precision of the map, and also gives users the flexibility to derive their own maps with a customised legend, rather than relying on the definition that the mapmakers provide. Thus exhaustive land cover fraction maps are desired by a number of user communities.

We compared the performance of eight machine learning regression algorithms for land cover fraction mapping. We used and assessed six types of covariates for training the models: spectral bands and vegetation indices derived from Proba-V 100 m TOC reflectance; temporal features from its time series; terrestrial features from ASTER GDEM; soil properties from SoilGrids; climate data from WorldClim; and location. The models were trained on 138164 training points throughout the globe, describing land cover fractions of pixels aligned to the Proba-V 100 m grid. After training, their land cover class fraction predictions were compared with an independent validation set of 20705 points, likewise aligned to the Proba-V grid.

One challenge for fraction mapping models is data sparsity: most fractions are 0\%, but regression favours the mean, thus 0\% and 100\% are difficult to predict well. We propose a new solution by combining three models. A binary model determines whether a pixel is pure; if so, it is processed using a classification model; otherwise with a regression model.

Results show that the three-model approach results in better estimation of extreme values (lower MAE of 8.4 vs 9.4) at the cost of misclassification (higher RMSE of 19.4 vs 17.3). From the compared models, random forest and cubist regressions reach the highest overall accuracy of 72\%, comparable to the reported accuracies of the current discrete global land cover maps. The models were driven primarily by spectral remote sensing data, followed by soil and climate data.

This research proves that machine learning algorithms can be applied globally to a wide variety of land cover to create an exhaustive global land cover fraction map. This expresses land cover more precisely, and empowers users to create their own discrete maps, which helps land cover monitoring and climate modelling communities. In addition, this work contributes to efforts to produce exhaustive global land cover fractions as an operational product.
